% -*- compile-command: "latexmk -pdf document.tex" -*-
\documentclass{article}

\author{Sam Price}
\title{Math 452: Assignment 4}

\newif\ifprinted%
\printedtrue%
\input{preamble}

\begin{document}

\maketitle

\begin{enumerate}[start=0,label={RELABEL}]
  \item[14.8\rparen] Explain why $\QQ$ is not cyclic.

        The rationals are not cyclic because there is no $p/q$ such that for \emph{every} $s/t$ there is an integer $k$ so that $kp/q = s/t$.
        For instance, since we take $(p, q)$ to be coprime, simply take $(p, q^{2})$. If $q = 1$, then of course an integer may not generate all rationals.

  \item[14.15\rparen] Find all generators of $\ZZ_{3} \times \ZZ_{4}$.

        The generators are
        \[ (1, 2), (2, 1), (1, 3), (2, 3). \]

  \item[\textbf{14.21}\rparen] Let $H \le G$ and $g \in G$ fixed.
        Prove that if $G$ is cyclic, then the conjugate of $H$ w.r.t. $g$ is cyclic.

        \begin{proof}
          Since $gHg\inv$ is a subgroup of $G$ and $G$ itself is cyclic,
          we know $gHg\inv$ must be cyclic as well.
        \end{proof}

  \item[16.12\rparen] Let $G$ be a group and fix $g \in G$.
        Define $\lambda \from G \to G$ so that $\lambda(a) = gag\inv$. Prove $\lambda$ is an isomorphism.

        \begin{proof}
          Note that $\lambda(ab) = gabg\inv = ga (gg\inv) bg = \lambda(a)\lambda(b)$.
          Since $\lambda$ is a homomorphism, we need only show that $\ker \lambda$ is trivial now,
          as then $\lambda$ would be injective and so surjective.
          Suppose there is some $a \in G$ so that $\lambda(a) = e$.
          Then, $gag\inv = e$ which implies that $ga = g$ and so $a = e$.
        \end{proof}

  \item[\textbf{16.16}\rparen] Determine if each pair of groups is isomorphic.

        \begin{lemma}{}{zn-zm-iso-cyc}
          Let $A, B$ be cyclic groups of order $m, n$ respectively.
          Then, $A \times B$ is cyclic if and only if $\gcd(m, n) = 1$.
        \end{lemma}
        \begin{proof}
          ($\impliedby$)
          Suppose $\gcd(m, n) = 1$ and $a, b$ be generators of $A$ and $B$ respectively.
          $A \times B$ is cyclic only if there is an element of order $mn$ within it.
          We claim that $(a, b)$ is such an element.
          Clearly, its order is some divisor of $mn$, that much is clear by simple inspection.
          Suppose its order is $k < mn$. This means that $a^{k} = b^{k} = e$,
          but then $m \mid k$ and $n \mid k$ which is a contradiction as $\lcm(m, n) = mn$.% Comment for footnote spacing
          \footnote{Yes, I get the abuse of notation here by not distinguishing between $e_{A}, e_{B}$ and $e_{A \times B}$.}
          Hence, $(a, b)$ has order $mn$ and so $A \times B$ is cyclic.

          ($\implies$)
          Suppose now $\gcd(m, n) > 1$. Let $a^{k}, b^{j}$ be any elements of $A, B$ for generators $a, b$.
          Thus $\lcm(m, n) < mn$ and so
          \[ \pars{a^{k}, b^{j}}^{\lcm(m, n)} = e. \]
          Since every element has order less than $mn$, $A \times B$ is not cyclic.
        \end{proof}


        \begin{enumerate}[start=1,label={(\alph*)}]
          \item Yes.

          \item No.

          \item No. First is not cyclic, second one is.
        \end{enumerate}

        Addendum: The reason behind the one-word answers is that if $A$ and $B$ are both cyclic groups, and $\abs{A} = \abs{B}$,
        then it is trivial to show that $A \iso B$.
        Sketch: Send generator powers to each other.

\newpage% Formatting

  \item[16.19\rparen] Let $\theta \from \ZZ_{12} \to \ZZ_{12}$ defined by $n \mapsto 4n$.

        \begin{itemize}
          \item Verify $\theta$ is operation-preserving.

                Let $a, b \in \ZZ_{12}$. Then,
                \[ \theta(a + b) = 4(a + b) = 4a + 4b = \theta(a) + \theta(b). \]

          \item Is $\theta$ an isomorphism?

                No, as $\ker \theta$ is nontrivial. Note $3 \in \ker \theta$, which is enough.
        \end{itemize}


  \item[17.6\rparen] Let $f \from \ZZ_{15} \to \ZZ_{20}$ be defined by $n \mapsto 4n$. Show $f$ is a homomorphism.

        Firstly, $f(0) = 0$. Let $a, b \in \ZZ_{15}$.
        We find that $f(a + b) = 4(a + b) = 4a + 4b = f(a) + f(b)$.
        We don't bother checking if they are equal when crossing modulo-indices since we can take equivalence classes and addition is fine across those.
        That is, $[a + b] = [a] + [b]$.

  \item[17.7\rparen] For the same $f$ as in the previous question:

        \begin{enumerate}[start=1,label={(\alph*)}]
          \item Find $\ord(7)$ and $\ord(f(7))$.

                Easily $\ord(7) = 15$ since they are coprime.
                Then, we know $\ord(f(7)) \mid 15$.
                In fact, we know with this specific $f$ every order of $f(n)$ must divide 5. So, since it is not in $\ker f$, it is 5 exactly.

          \item Same for 6.

                Order 5 for both. See reasoning for the latter above.

          \item Run it back for 10.

                $\ord(10) = 3$ and has order 0 under $f$ since it is in the kernel of $f$.

          \item Is this surprising? Explain.

                Not really. The $n \mapsto 4n$ map really trivializes it as either the order of something under $f$ is 0 (when $a \in \ker f$)
                or must be 5 itself since $20/4 = 5$ is prime.
        \end{enumerate}


  \item[17.10\rparen] Consider $\lambda \from U_{13} \to U_{13}$ given by $a \mapsto a^{3}$.

        \begin{enumerate}[start=1,label={(\alph*)}]
          \item How are 2 and 7 related in the domain of $U_{13}$?

                They are related by the fact $7 \cdot 2 = 1$.

          \item Their images?

                In this case $\lambda(7) \lambda(2) = 7^{3}2^{3} = 1 = \lambda(2 \cdot 7)$.

          \item Repeat for elements $4, 10 \in U_{13}$.

                We see $4 \cdot 10 = 1$. Their images hold that $\lambda(4)\lambda(10) = 4^{3}10^{3} = 1 = \lambda(4 \cdot 10)$.

          \item Is this surprising? Explain.

                No, since homomorphisms must preserve inverses and group operations.
        \end{enumerate}

\end{enumerate}


\end{document}
